\plaintitle

### 2016-09-22 drafts/labor/complexity.tex
internet, not Internet
complexity tweaks

\topic{Crowdsourcing research has spent the better part of a decade
proving the viability of crowdsourcing in complex work.}
Unless crowdsourcing can demonstrate viability for meaningfully complex tasks,
the argument runs,
it will be incapable of ensuring a pro--social outcome for work and workers~\cite{crowdworkFuture}.
\citeauthor{crowdForgeKittur} first opened the question of
whether crowdsourcing could be used for goals that are not simple parallel tasks~\cite{crowdForgeKittur}.
Their work demonstrated proof--of--concept crowdsourcing of
a simple encyclopedia article and news summary
--- tasks which could be verified or repeated
with reasonable expectations of similar outcomes.
Seeking to raise the complexity ceiling~\cite{myers2000past},
researchers have since created
additional proof--of--concept applications and techniques,
including conversational assistants \cite{Lasecki:2013:CCC:2501988.2502057},
medical data interpreters \cite{Lasecki:2013:CCC:2501988.2502057}, and
idea generation \cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
to name a few examples.

@@

It is now clear that this approach works with focused complex tasks, but
the broader wicked problem that each complex goal represents remains unsolved~\cite{rittel1973dilemmas}.
As a first example,
idea generation shows promise~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
but there is as yet no general crowdsourced solution for
the broader goal of invention and innovation~\cite{fuge2014analysis}.
Second,
focused writing tasks are now feasible~\cite{Kim2017,bernsteinSoylent,Nebeling:2016:WCW:2858036.2858169,
      writingMicroTasks,agapie2015crowdsourcing}, but
there is no general solution to create
a cross--domain, high--quality crowd--powered author. 
Third,
data analysis tasks such as
clustering \cite{chilton2013cascade},
categorization \cite{andre2014crowd}, and
outlining \cite{luther2015crowdlines}
are possible, but there is no general solution for sensemaking.
It is not yet clear what insights would be required
to enable crowdsourced solutions for these broader wicked problems.
%  and
% other crowdsourced goals such as addressing climate change
% \cite{introne2011climate}.

@@

% \topic{Digital media have expanded the scope of
% viable piecework by pushing drastically on the limits
% cited by piecework researchers.}
The research on piecework tells us that
we should expect piecework to thrive in industries where
the nature of the work is limited in complexity~\cite{Brown01041990}.
Given the flourishing of on--demand labor platforms such as
Uber, AMT, and others, we ask ourselves
what --- if anything --- has changed.
We argue that
the Internet has trivialized
the costs and challenges of the earlier limiting factors
because technology makes it easier
\begin{inlinelist}
\item for workers to do complex work without training,
\item to manage workers in doing complex work, and 
\item to create the infrastructure necessary to manage the workers.
\end{inlinelist}

@@

\subsubsection{\implication}
% \topic{The piecework literature gives us a template for pushing the boundaries of complexity in piecework, but
% it also signals some of the ultimate limitations of crowd work and piecework in general.}
% While the threshold preventing task requesters from utilizing piecework
% has dropped thanks to affordances of the Internet, the ceiling on task complexity hasn't moved significantly.
% \msb{is that your prediction? I would argue against the fact that it hasn't moved significantly, crowdforge did far more complex work, as did flash teams and flash orgs}
% If we're to make use of \citeauthor{Brown01041990}'s prescriptions,
% we would benefit from finding ways to decompose varied tasks into homogeneous microtasks. \msb{isn't that what we've been doing all along?}
Technology's ability to support human cognition will enable stronger assumptions about workers' abilities, increasing the complexity of crowd work outcomes.
Just as the shift to expert crowdsourcing increased complexity, so too will workers with better tools increase the set of tasks possible.
Beyond this, further improvements would most likely come from replicating the success of narrowly--slicing education for expert work as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
of human computation \cite{grier2013computers} and drastically reformulating macro--tasks given the constraints of piecework \cite{hart2013rise}.
To some extent, an argument can be made that
MOOCs and other online education resources
provide crowd workers with the resources that they need, but 
it remains to be seen whether that work will be appropriately valued, let alone
properly interpreted by task solicitors~\cite{aguaded2013mooc}.
If we can overcome this obstacle,
we might be able to empower more crowd workers to do complex work such as engineering and metalworking,
rather than doom them to ``uneducated'' match girl reputations~\cite{10.2307/3827491}.
However, many such experts are already available on platforms such as Upwork, so training may not directly increase the complexity accessible to crowd work unless it makes common expertise more broadly available.

>>>>

\topic{Crowdsourcing research has spent the better part of a decade
proving the viability of crowdsourcing in complex work.}
Unless crowdsourcing can demonstrate viability for meaningfully complex tasks,
the argument runs,
it will be incapable of ensuring a pro--social outcome for work and workers~\cite{crowdworkFuture}.
\citeauthor{crowdForgeKittur} first opened the question of
whether crowdsourcing could be used for goals that are not simple parallel tasks~\cite{crowdForgeKittur}.
Their work demonstrated proof--of--concept crowdsourcing of
a simple encyclopedia article and news summary
--- tasks which could be verified or repeated
with reasonable expectations of similar outcomes.
Seeking to raise the complexity ceiling~\cite{myers2000past},
researchers have since created
additional proof--of--concept applications and techniques,
including conversational assistants \cite{Lasecki:2013:CCC:2501988.2502057},
medical data interpreters \cite{Lasecki:2013:CCC:2501988.2502057}, and
idea generation workflows \cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
to name a few examples.

@@

It is now clear that this computational workflow approach works with focused complex tasks, but
the broader wicked problems largely remain unsolved~\cite{rittel1973dilemmas}.
As a first example,
idea generation shows promise~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
but there is as yet no general crowdsourced solution for
the broader goal of invention and innovation~\cite{fuge2014analysis}.
Second,
focused writing tasks are now feasible~\cite{Kim2017,bernsteinSoylent,Nebeling:2016:WCW:2858036.2858169,
      writingMicroTasks,agapie2015crowdsourcing}, but
there is no general solution to create
a cross--domain, high--quality crowd--powered author. 
Third,
data analysis tasks such as
clustering \cite{chilton2013cascade},
categorization \cite{andre2014crowd}, and
outlining \cite{luther2015crowdlines}
are possible, but there is no general solution for sensemaking.
It is not yet clear what insights would be required
to enable crowdsourced solutions for these broader wicked problems.
%  and
% other crowdsourced goals such as addressing climate change
% \cite{introne2011climate}.

@@

% \topic{Digital media have expanded the scope of
% viable piecework by pushing drastically on the limits
% cited by piecework researchers.}
The research on piecework tells us that
we should expect piecework to thrive in industries where
the nature of the work is limited in complexity~\cite{Brown01041990}.
Given the flourishing of on--demand labor platforms such as
Uber, AMT, and others, we ask ourselves
what --- if anything --- has changed.
We argue that
the internet has trivialized
the costs and challenges of the earlier limiting factors
because technology makes it easier
\begin{inlinelist}
\item for workers to do complex work without training,
\item to manage workers in doing complex work, and 
\item to create the infrastructure necessary to manage the workers.
\end{inlinelist}

@@

\subsubsection{\implication}
% \topic{The piecework literature gives us a template for pushing the boundaries of complexity in piecework, but
% it also signals some of the ultimate limitations of crowd work and piecework in general.}
% While the threshold preventing task requesters from utilizing piecework
% has dropped thanks to affordances of the internet, the ceiling on task complexity hasn't moved significantly.
% \msb{is that your prediction? I would argue against the fact that it hasn't moved significantly, crowdforge did far more complex work, as did flash teams and flash orgs}
% If we're to make use of \citeauthor{Brown01041990}'s prescriptions,
% we would benefit from finding ways to decompose varied tasks into homogeneous microtasks. \msb{isn't that what we've been doing all along?}
Technology's ability to support human cognition will enable stronger assumptions about workers' abilities, increasing the complexity of crowd work outcomes.
Just as the shift to expert crowdsourcing increased complexity, so too will workers with better tools increase the set of tasks possible.
Beyond this, further improvements would most likely come from replicating the success of narrowly--slicing education for expert work as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
of human computation \cite{grier2013computers} and drastically reformulating macro--tasks given the constraints of piecework \cite{hart2013rise}.
To some extent, an argument can be made that
MOOCs and other online education resources
provide crowd workers with the resources that they need, but 
it remains to be seen whether that work will be appropriately valued, let alone
properly interpreted by task solicitors~\cite{aguaded2013mooc}.
If we can overcome this obstacle,
we might be able to empower more crowd workers to do complex work such as engineering and metalworking,
rather than doom them to ``uneducated'' match girl reputations~\cite{10.2307/3827491}.
However, many such experts are already available on platforms such as Upwork, so training may not directly increase the complexity accessible to crowd work unless it makes common expertise more broadly available.
